
---
title: "Bootstrap using Infer package"
author: "Rigoberto Leyva, Xiaodan Chen, Yijie Ying"
date: "2/10/2021"
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "dolphin"
    fonttheme: "structurebold"

---


## What is the infer package?

-   Package to perform statistical inference based on tidyverse design.

-   Four main verbs:

    -   \textbf{\textsl{specify()}} specifies the y variable or a model with y (response) and x (explanatory) variables.

    -   \textbf{\textsl{hypothesize()}} declare the null hypothesis.

    -   \textbf{\textsl{generate()}} generate data.

    -   \textbf{\textsl{calculate()}} calculates a distribution of statistics from the data.

## Specify() use

\begin{center}

  \textcolor{purple}{\textbf{\textsl{specify(x, formula, response = NULL, explanatory = NULL, \\
                          success = NULL)}}}

\end{center}

-   Specifying the y variable:

    -   Gestation_sample %\>%\
        specify(response = gestation) or specify(response = gestation, explanatory = age)

-   To specify a formula such as regression.

    -   Gestation_sample %\>%\
        specify(formula = gestation \~ age)

## Hypothesize() use

\begin{center}

  \textcolor{purple}{\textbf{\textsl{hypothesize(x, null, p = NULL, mu = NULL, med = NULL, \\ 
  sigma = NULL)}}}

\end{center}

-   Setting null hypothesis $H_o: \mu = 250$:

    -   Gestation_sample %\>%\
        specify(response = gestation) %\>%\
        hypothesize(null = "point", mu = 25)

## Generate() use

\begin{center}
  
  \textcolor{purple}{\textbf{\textsl{generate(x, reps = 1, type = NULL, ...)}}}

\end{center}

-   Generate 200 bootstrap samples:

    -   Gestation_sample %\>%\
        specify(response = gestation) %\>%\
        hypothesize(null = "point", mu = 250) %\>%\
        generate(reps = 200, type = "bootstrap")

## Calculate() use

\begin{center}

  \textcolor{purple}{\textbf{\textsl{calculate(x,
      stat = c("mean", "median", "sum", "sd", "prop", "count", 
               "diff in means", "diff in medians", "diff in props", 
               "Chisq", "F", "slope", "correlation", "t", "z", 
               "ratio of props", "odds ratio"),
      order = NULL,
      ...
    )}}}
    
\end{center}

-   Calculate summary statistic for the mean.

    -   Gestation_sample %\>%\
        specify(response = gestation) %\>%\
        hypothesize(null = "point", mu = 250) %\>%\
        generate(reps = 200, type = "bootstrap") %\>%\
        calculate(stat = "mean")

## Reference & Example 

\textcolor{purple}{\textbf{Reference:}}

• Ismay, Kim, \textbf{ModernDive}    
\textsl{(https://moderndive.com/index.html)}    
• Baumer, Kaplan, Horton, \textbf{Modern Data Science with R}, 2nd edition  \textsl{(https://mdsr-book.github.io/mdsr2e/index.html)}


\textcolor{purple}{\textbf{Example:}}

Calculate and interpret a 95% confidence interval for the mean age of mothers from the Gestation data set from the mosaicData package.

```{r}
library(pacman)
p_load(infer, mosaicData, tidyverse)
```

## Data Set Structure

```{r, echo = FALSE}
Gestation <- Gestation %>% filter(!is.na(age))
```


\tiny   
```{r cars, echo = TRUE}
glimpse(Gestation)
```

## **Specify()** - Select Variable

\small 
```{r,message=FALSE}
Gestation %>%
  specify(response = age)
```

## **Generate()** - Draw Bootstrap Samples

\small 
```{r}
Gestation %>%
  specify(response = age) %>%
  generate(reps = 100, type = 'bootstrap')
```

## **Calculate()** - Evaluate Bootstrap Replications

\small 
```{r}
Gestation %>%
  specify(response = age) %>%
  generate(reps = 100, type = 'bootstrap') %>%
  calculate(stat = 'mean') 
```

## **Visualize()** - Visualize the Bootstrap Distribution

\small 
```{r out.width = '60%', fig.align = "center"}
bootstrap_distribution <- Gestation %>%
  specify(response = age) %>%
  generate(reps = 100, type = 'bootstrap') %>%
  calculate(stat = 'mean') 
visualize(bootstrap_distribution)
```


## Get_confidence_interval() / get_ci() use

\begin{center}

  \textcolor{purple}{\textbf{\textsl{get\_confidence\_interval(x, level = 0.95, 
                         type = "percentile", point\_estimate = NULL)}}}
    
\end{center}


-   \footnotesize x : data frame of calculated statistics 
    
-   level : 0 to 1 ; default value is 0.95

-   type : method for creating the confidence interval

-   point_estimate : needed if type is "se" or "bias-corrected"


## 95% Confidence Intervals - Standard Error Method

$$\big(\bar x - 1.96\cdot se(\bar x), \bar x + 1.96\cdot se(\bar x)\big)$$

\small 
```{r,message=FALSE}
x_bar <- bootstrap_distribution %>% 
  summarise(x_bar <- mean(stat))
standard_error_ci <- bootstrap_distribution %>%
  get_confidence_interval(type = 'se', point_estimate = x_bar, 
                          level = 0.95)

standard_error_ci
```

## 95% Confidence Intervals - Standard Error Method

\small 
```{r out.width = '75%', fig.align = "center"}
visualize(bootstrap_distribution) +
  shade_confidence_interval(endpoints = standard_error_ci)
```

## 95% Confidence Intervals - Percentile Method

\small 
```{r}
percentile_ci <- bootstrap_distribution %>%
  get_confidence_interval(level = 0.95, type = 'percentile')
percentile_ci
```

## 95% Confidence Intervals - Percentile Method

\small 
```{r out.width = '70%', fig.align = "center"}
visualize(bootstrap_distribution) +
  shade_confidence_interval(endpoints = percentile_ci, 
                          color = 'lightblue', fill = 'lightblue')
```

## Reference

• Ismay, Kim, ModernDive    

\textsl{(https://moderndive.com/index.html)}    

• Baumer, Kaplan, Horton, Modern Data Science with R, 2nd edition  

\textsl{(https://mdsr-book.github.io/mdsr2e/index.html)}

• infer R Package

\textsl{https://infer.netlify.app/index.html}

• Package ‘infer’

\textsl{https://cran.r-project.org/web/packages/infer/infer.pdf}









